---
title: "Project"
author: "Ajay Rawtani"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(coop)
library(lsa)
library(reshape2)

```


```{r}
dat_obj <- read.csv("Store_Data.csv")
head(dat_obj)
```

```{r cleaning}
dat_obj <- dat_obj %>% filter(Qty>0)

dat_obj <- na.omit(dat_obj)

dat_obj <- subset(dat_obj, !tolower(Customer.Code) %in% c("him", "himw","91"," "))
dat_obj <- subset(dat_obj, !tolower(Customer.Code) %in% "919642306464")

write.csv(dat_obj,"Cleaned_Strore_Data.csv",row.names = FALSE)

dat_obj$Bill.Date =as.POSIXct(dat_obj$Bill.Date,
                            format="%m/%d/%Y")

dat_obj

# There seems to be many rows of NA values that we need to handle
# We can just omit these rows using na.omit
# cat("\nThe number of NA values in the dataset before cleaning is:",sum(is.na(dat_obj$Customer.Code)))
# 
# dat_obj <- na.omit(dat_obj)
# 
# cat("\nThe number of NA values in the dataset after cleaning is:",sum(is.na(dat_obj$Customer.Code)))

```
```{r}

# correcting the date field
Cleaned_Strore_Data <- read_csv("~/DU/Probability and Statistics 2/Project/Cleaned_Strore_Data.csv")
dat_obj <- Cleaned_Strore_Data
dat_obj
dat_obj$Bill.Date <- as.Date(dat_obj$Bill.Date, format = "%d/%m/%Y")
dat_obj
dat_obj$Bill.Date <- format(dat_obj$Bill.Date, "%Y-%m-%d")

#Create a unique bill item Number:
dat_obj$Bill.Item <- paste0(as.character(dat_obj$Bill.No),"_",as.character(dat_obj$Bill.Prefix))
dat_obj

#Saving document
write.csv(dat_obj, "Updated_Store_Data_Cleaned.csv", row.names = FALSE)
```

## Checking the trends

Plotting some of the data to gain insights before performing any sort of analysis on the product data

```{r}
#Reading file
dat_obj <- read.csv("Updated_Store_Data_Cleaned.csv")
dat_obj$Bill.Date <- as.Date(dat_obj$Bill.Date, format = "%Y-%m-%d")

# looking at the invoice data,i.e., the number of orders placed by customers over time
total_invoices <- dat_obj %>%
  group_by(Bill.Date=floor_date(Bill.Date, "month")) %>%
  summarise(number_of_orders=n_distinct(Bill.Item))

ggplot(total_invoices, aes(x = Bill.Date, y = number_of_orders)) +
  geom_line(color = "steelblue") +
  labs(x = "Invoice Date", y = "Number of Orders") +
  ggtitle("Number of orders placed over a period") +
  theme_bw() + scale_x_date(date_breaks = "3 month")

# The dip in the plot is due to insufficient data in the month of December(only 9 days of data was collected/available)
revenue_over_time <- dat_obj %>%
  group_by(Bill.Date=floor_date(Bill.Date, "month")) %>%
  summarise(Sales=sum(Value))

# revenue_over_time
ggplot(revenue_over_time, aes(x= Bill.Date, y=Sales)) + geom_line(color = "red")+ labs(x= "Invoice Date", y = "Sales") + ggtitle("Revenue of the company over a period")+theme_bw() + scale_x_date(date_breaks = "3 month")

# Average revenue per bill.items grouped by bill.time
average_revenue <- dat_obj %>%
  group_by(Bill.Date=floor_date(Bill.Date, "month")) %>%
  summarise(average_revenue = sum(Value) / n_distinct(Bill.Item))

# Plot of Average revenue per Bill.Item over time
ggplot(average_revenue, aes(x = Bill.Date, y = average_revenue)) +
  geom_line(color = "green") +
  labs(x = "Invoice Date", y = "Average Revenue per Bill.Item") +
  ggtitle("Average Revenue per Bill.Item over Time") +
  theme_bw() +
  scale_x_date(date_breaks = "3 month")
```

## Repeat Customers

Lets find out if there are repeat customers in this dataset. We have the customer ID of the purchases. We can group by and check out what kind of analysis can be performed   

```{r}
#aggregating data so that one row represents one purchase order
Invoice <- dat_obj %>%
  group_by(Bill.Date, Bill.Item) %>%
  summarise(total_sales = sum(Value), total_quantity = sum(Qty),Customer.Code = max(Customer.Code))

Invoice
#aggregating data into months
InvoiceCustomer <- 
  Invoice %>% 
  group_by(Bill.Date = floor_date(Bill.Date, "month"), Customer.Code) %>%
  summarise(Count = n_distinct(Bill.Item), Sales = sum(total_sales))
InvoiceCustomer
write.csv(InvoiceCustomer,"InvoiceCustomer.csv", row.names = FALSE)

# Repeat Customers are count of unique customer.code grouped by date
RepeatCustomers <- InvoiceCustomer %>%
  filter(Count > 1)

RepeatCustomers <- RepeatCustomers %>%
  group_by(Bill.Date) %>%
  summarize(Count=n_distinct(Customer.Code), Sales=sum(Sales))
RepeatCustomers

#total number of monthly customers
UniqueCustomers <- dat_obj %>%
  group_by(Bill.Date=floor_date(Bill.Date, "month")) %>%
  summarise(Count=n_distinct(Customer.Code))

#find the percentage of monthly revenue that are attributed to the repeat customers
RepeatCustomers$Perc <- RepeatCustomers$Sales/revenue_over_time$Sales*100.0
RepeatCustomers$Perc
#append unique customers
RepeatCustomers$Total <- UniqueCustomers$Count
RepeatCustomers$Total

ggplot(RepeatCustomers) +
  geom_line(aes(x=Bill.Date, y=Total, color="Count of Total Unique Customers"), stat="identity") +
  geom_line(aes(x=Bill.Date, y=Count, color="Count of Total Repeat Customers"), stat="identity") +
  # geom_bar(aes(x=Bill.Date, y=Perc*20, fill="Percentage of Revenue"), stat="identity", alpha=0.5) +
  # scale_color_manual(values=c("orange", "navy"), labels=c("Total Unique Customers", "Repeat Customers")) +
  # scale_fill_manual(values="gray", labels="Revenue from repeat customers") +
  # scale_y_continuous(sec.axis = sec_axis(~./20, name="Percentage (%)")) +
  labs(title="Number of Unique vs Repeat Customers") +
  theme_bw() +
  theme(legend.position="top")

#Revenue from repeat customer
RepeatCustomers$RevenueRepeatCustomers <- RepeatCustomers$Sales * RepeatCustomers$Count / RepeatCustomers$Total

# Plot the line chart
ggplot(RepeatCustomers) +
  geom_line(aes(x = Bill.Date, y = RevenueRepeatCustomers, color = "Revenue from Repeat Customers"), stat = "identity") +
  labs(title = "Number of Unique vs Repeat Customers and Revenue from Repeat Customers") +
  theme_bw() +
  theme(legend.position = "top")
```
## Top-sellers

```{r}
# Number of items sold for each product
pop.products <- dat_obj %>%
  group_by(StockNo) %>%
  summarise(Quantity = sum(Qty))

# Rank products based on total quantity sold
top.products <- pop.products %>%
  arrange(desc(Quantity)) %>%
  top_n(5)


# Retrieve stock names
stock_names <- dat_obj %>%
  filter(StockNo %in% top.products$StockNo) %>%
  distinct(StockNo, Item.Description)

# Merge stock names with top.products
top.products <- top.products %>%
  left_join(stock_names, by = "StockNo")


# Plot the best selling products
ggplot(top.products, aes(x = Item.Description, y = Quantity, fill = StockNo)) +
  geom_bar(stat = "identity") +
  labs(title = "Best Selling Products of All Time",
       x = "Product",
       y = "Quantity Sold") +
  theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
cust.item.mat<- dcast(dat_obj, Customer.Code ~ StockNo, value.var = "Qty")

purchase.check <- function(x){
  as.integer(x>0)
}

cust.item.mat <- cust.item.mat %>% mutate_at(vars(-Customer.Code), funs(purchase.check))
cust.item.mat

```

## User based collaborative filtering


```{r}
#Calculate the cosine
usercosine <- cosine(as.matrix(t(cust.item.mat[, 2:dim(cust.item.mat)[2]])))
colnames(usercosine) <- cust.item.mat$Customer.Code
```


```{r}
user_number <- readline("\nEnter a valid number: ")
#rank the most similar customers to our customer with Customer Code
Top5Similar <- cust.item.mat$Customer.Code[
  order(usercosine[, user_number], decreasing = TRUE)[2:6]]

# Similar Users
cat('\nTop 5 similar users:', Top5Similar)

#Pick one user (temp fix)
sim_usr <- sample(Top5Similar, 1)
cat('\n Picking a similar user:', sim_usr)

#let's find what the customer A(12350) bought
boughtbyA <- cust.item.mat %>%
  filter(Customer.Code == user_number)
boughtbyA <- colnames(cust.item.mat)[which(boughtbyA !=0)]
cat('\nItem bought by select user', user_number,':\n')
boughtbyA[-1]

#let's find the descriptions of these items
boughtbyADescription <- unique(dat_obj[which(dat_obj$StockNo %in% boughtbyA),
     c("StockNo", "Item.Description")])
boughtbyADescription

#let's find what bought the B customer (I pick a customer with 
#ID=17935 from Top10Similar)
boughtbyB <- cust.item.mat %>%
  filter(Customer.Code == sim_usr)
boughtbyB <- colnames(cust.item.mat)[which(boughtbyB !=0)]
cat('\nItem bought by similar user', user_number,':')
boughtbyB[-1]

#let's find the items that the customer B didn't buy
#so we can recommend these items to buy for B
RecommendToA <-setdiff(boughtbyB,boughtbyA)
cat("\nProduct Recommendations:", RecommendToA)

#let's find the descriptions of these items
RecommendToADescription <- unique(
  dat_obj[which(dat_obj$StockNo %in% RecommendToA),
     c("StockNo", "Item.Description")])

RecommendToADescription <- RecommendToADescription[match(RecommendToA, RecommendToADescription$StockNo),
]
#here is the list of the items descriptions as a recommendation to B
cat('Product Recommended for purchase/marketing to A:')
RecommendToADescription

```

## Item-based Recommendation

```{r}
itemcosine <- cosine(as.matrix(cust.item.mat[, 2:dim(cust.item.mat)[2]]))
```

```{r}
# Read Stock Code
usr_stockcode <- readline('Enter stock number: ')
#get Stock Code descriptions
StockDescriptions <- unique(
    dat_obj[which(dat_obj$StockNo %in% usr_stockcode), c("StockNo", "Item.Description")])
StockDescriptions

#find top5 most similar products to the product with Stock Code
Top5SimilarItems <- colnames(itemcosine)[
    order(itemcosine[, usr_stockcode], decreasing = TRUE)[2:6]]

#get descriptions
Top5SimilarItemsDescriptions <- unique(
    dat_obj[which(dat_obj$StockNo %in% Top5SimilarItems), c("StockNo", "Item.Description")])

Top5SimilarItemsDescriptions <- Top5SimilarItemsDescriptions[
    match(Top5SimilarItems, Top5SimilarItemsDescriptions$StockNo),]

Top5SimilarItemsDescriptions
```

